# Plotly R graphing

Brian Mao

## Introduction

Plotly's R graphing library makes interactive, publication-quality graphs. in this into to Plotly I will show some examples of how to make line plots, scatter plots, area charts, bar charts, box plots, histograms, heatmaps, and 3D scatter plot charts. there are a lots more advanced function in the Plotly. you may explore it later on yourself 

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## URL

URL https://plotly-r.com, https://github.com/ropensci/plotly#readme,
https://plotly.com/r/

## Install and packages

```{r message=FALSE}
library(plotly)
library(ggplot2)
```

## Basic examples

### Basic Scatter Plot

```{r message=FALSE}
library(plotly)
fig <- plot_ly(data = iris, x = ~Sepal.Length, y = ~Petal.Length)
fig
```

### Plotting Markers and Lines

```{r message=FALSE}
library(plotly)
trace_0 <- rnorm(100, mean = 5)
trace_1 <- rnorm(100, mean = 0)
trace_2 <- rnorm(100, mean = -5)
x <- c(1:100)
data <- data.frame(x, trace_0, trace_1, trace_2)
fig <- plot_ly(data, x = ~x)
fig <- fig %>% add_trace(y = ~trace_0, name = 'trace 0',mode = 'lines')
fig <- fig %>% add_trace(y = ~trace_1, name = 'trace 1', mode = 'lines+markers')
fig <- fig %>% add_trace(y = ~trace_2, name = 'trace 2', mode = 'markers')
fig
```


### Data with different Symbols

```{r message=FALSE}
library(plotly)
fig <- plot_ly(data = iris, x = ~Sepal.Length, y = ~Petal.Length, type = 'scatter',
  mode = 'markers', symbol = ~Species, symbols = c('circle','x','o'),
  color = I('black'), marker = list(size = 10))
fig
```

## Statistical Charts

### Basic Boxplot

```{r message=FALSE}
library(plotly)
fig <- plot_ly(y = ~rnorm(100), type = "box")
fig <- fig %>% add_trace(y = ~rnorm(300, 1))
fig
```


### overlaid Histograms

```{r message=FALSE}
fig <- plot_ly(alpha = 0.6)
fig <- fig %>% add_histogram(x = ~rnorm(1000))
fig <- fig %>% add_histogram(x = ~rnorm(500) + 1)
fig <- fig %>% layout(barmode = "overlay")
fig
```


### Basic Heatmap

```{r message=FALSE}
library(plotly)
fig <- plot_ly(z = volcano, type = "heatmap")
fig
```

## more advanced map plot

### Flight Paths Map

```{r message=FALSE}
library(plotly)

fig <- plot_geo(lat = c(40.7127, 51.5072), lon = c(-74.0059, 0.1275))
fig <- fig %>% add_lines(color = I("blue"), size = I(2))
fig <- fig %>% layout(
    title = 'London to NYC Great Circle',
    showlegend = FALSE,
    geo = list(
      resolution = 50,
      showland = TRUE,
      showlakes = TRUE,
      landcolor = toRGB("grey80"),
      countrycolor = toRGB("grey80"),
      lakecolor = toRGB("white"),
      projection = list(type = "equirectangular"),
      coastlinewidth = 2,
      lataxis = list(
        range = c(20, 60),
        showgrid = TRUE,
        tickmode = "linear",
        dtick = 10
      ),
      lonaxis = list(
        range = c(-100, 20),
        showgrid = TRUE,
        tickmode = "linear",
        dtick = 20
      )
    )
  )

fig
```


### 3D Scatter Plot

```{r message=FALSE}
library(plotly)

mtcars$am[which(mtcars$am == 0)] <- 'Automatic'
mtcars$am[which(mtcars$am == 1)] <- 'Manual'
mtcars$am <- as.factor(mtcars$am)

fig <- plot_ly(mtcars, x = ~wt, y = ~hp, z = ~qsec, color = ~am, colors = c('#BF382A', '#0C4B8E'))
fig <- fig %>% add_markers()
fig <- fig %>% layout(scene = list(xaxis = list(title = 'Weight'),
                     yaxis = list(title = 'Gross horsepower'),
                     zaxis = list(title = '1/4 mile time')))

fig
```

## Conclusion

In this project I explored different methods of plotting interactive plot by utilizing plotly and ggplot2. you can explore more about the tool using R, python and Javascript. there are more resource available below. it is a very user friendly tool and the interface is well designed. it  allows users to design their own way to tell the story to its audience. effectively utilizing this tool will tremendously help you to improve your data visualization skills

## Work Cited:

Create Interactive Web Graphics via 'plotly.js'
https://cran.r-project.org/web/packages/plotly/plotly.pdf
Author Carson Sievert [aut, cre] (<https://orcid.org/0000-0002-4958-2844>),
Chris Parmer [aut],
Toby Hocking [aut],
Scott Chamberlain [aut],
Karthik Ram [aut],
Marianne Corvellec [aut] (<https://orcid.org/0000-0002-1994-3581>),
Pedro Despouy [aut],

https://github.com/plotly
https://plotly.com/graphing-libraries/






















Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
