# Introduction to Quandl()

Aiqi Zhou (az2638)

```{r , include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



Quandl::Quandl() 

## INTRODUCTION 

Quandl is a source of financial, economic and alternative data. It is acquired by Nasdaq in 2018 and has 20+ million data sets available. Users can acquire free data, purchase data or sell data at Quandl.

You can find data on a large variety of types of data from company data to the demographics data of a country. 
More can be explored on Quandl's website. https://www.quandl.com/search

The data is readily available in json, csv, xml formats and can be loaded in MatLab, R, Python, etc. 

All the data can be loaded in R through the package Quandl.


### Installation :

- install.packages("Quandl")

### Load Package 
- library(Quandl)

```{r}
library(Quandl)
```

### Set API key 

Quandl allows 50 calls a day for anonymous users.
Sign up for a free account to make unlimited calls every day.

Set up your own account and copy your api_key into the following function 

Quandl.api_key('your_key')

```{r, include=FALSE}
Quandl.api_key('JrvHzZcnakybdzAUChjn')
```



### Loading Data 

All available data can be found and viewed on https://www.quandl.com/search.
This package includes many free datasets, however, some datasets requires an account or is for paid use only.

To call on a data set in R. You must find the Quandl code of the data set. The codes can be found in the documentation page of the data set.

#### For example 

The follow dataset is called the "Federal Reserve Economic Data", which has the code FRED. 
This dataset has 335,000+ time-series data.

To call on certain aspect like the GDP of this data set, the Quandl code is "FRED/GDP".

```{r cars}

mydata = Quandl("FRED/GDP")
head(mydata)

```
To call on another aspect like disposable personal income, the Quandl code is "FRED/DSPI"

```{r}
mydata= Quandl("FRED/DSPI")
head(mydata)

```

### Type

The data can be called in different formats.
The default of the data is data frame, which can have an argument type = "raw"

- "raw"
- "ts"
- "zoo"
- "xts"
- "timeSeries"


```{r}
mydata = Quandl("FRED/GDP", type="xts")
head(mydata)
```
```{r}
mydata = Quandl("FRED/GDP", type="ts")
head(mydata)

```

```{r}
mydata = Quandl("FRED/GDP", type="zoo")
head(mydata)

```

### Transform 

There are some preprocessing can be done when loading the data. By adding a transform argument in Quandl(), the data can have the following types of transformations.

- "" = Default original data
- "diff" = row on row change in value
- "rdiff" = row on row percentage change
- "normalize" = scale to start at 100
- "cumul" = cumulative sum
- "rdiff_from" = latest value as % increment 

#### For example the percentage change of GDP by year

```{r}
mydata = Quandl("FRED/GDP",transform = "rdiff" )
head(mydata)
```

### Order 

The data can be ordered by date upon it is loaded.
Default is descending order by date.
- "desc"
- "asc"

```{r}
mydata = Quandl("FRED/GDP", order = "asc")
head(mydata)

```
### collapse

Since all data have a time associated. Quandl allows you to preprocess the data by specifying the frequency of the data.

- ''
- 'daily'
- 'weekly'
- 'monthly'
- 'quarterly'
- 'annual'

```{r}
mydata = Quandl("FRED/GDP", collapse = "quarterly")
head(mydata)
```




### Slicing data 

#### Rows
Because the data we are using is time series data, we can specify the range of row of the data set, using the start and end date arguments.

```{r}
mydata = Quandl("FRED/GDP", start_date="2001-12-31", end_date="2005-12-31")
head(mydata)
```

#### Columns
To get multiple columns of data from Quandl, the data can be called in the format: Quandl(c("col_1", "col_2"))

```{r}
mydata = Quandl(c("FRED/GDP", "FRED/DSPI"),start_date="2001-12-31", end_date="2005-12-31")
head(mydata)
```



### Finding a dataset in Quandl 

You can look for certain data sets within R using Quandl.search()

Format: 

Quandl.search(query = "search keyword", 
              page = # , 
              source = "source to search from if known", 
              silent = TRUE/FALSE)
              
- query: mandatory argument that you want to search for
- page: which page of search result you want. Default page = 1.
- source: specific source you want to search from.
- silent: print the results when FALSE. Nothing prints when true.

#### Example search for Japan 
```{r}
Quandl.search(query = "Japan", 
              page = 1 , 
              silent = FALSE)
```
You can pick which data set to use from the search result.

### Example 

The following is a quick example of getting data from the data set United Nations Industrial Commodities. 
From the documentation of this data set, we can learn what we are interested in. https://www.quandl.com/data/UINC-United-Nations-Industrial-Commodities

Here I want to plot the time series data of Beer production of Germany. The code for this information is "UINC/BEER_DEU".
I want to plot the time series, then the data need to be in the form of "ts"

```{r}
beer = Quandl("UINC/BEER_DEU",type = "ts")
head(beer)
```

```{r}
beer = Quandl("UINC/BEER_DEU",type = "ts")
plot.ts(beer)
```
If I want to see the yearly percentage change of the data, we need to add a transform argument "rdiff"
```{r}
beer_per_change = Quandl("UINC/BEER_DEU",type = "ts",transform = 'rdiff')
plot.ts(beer_per_change)

```




