# GGvis cheatsheet

Chenhui Mao

```{r, include=FALSE}
# keep this chunk in your .Rmd file
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

[Link to the cheatsheet](resources/ggvis_cheatsheet/ggvis_cheatsheet.pdf)

## ggvis

Official website: <http://ggvis.rstudio.com/>

ggvis is a visualization package that can:
- plot the graph in the same way as ggplot2
- create interaction strategies where you can interact with the graph locally or in the website

### Getting start

First thing first, we shall import some necessary packages

```{r}
#install.packages("dplyr")
#install.packages("ggvis")

library(ggvis)
library(dplyr)
```


This is a simple demonstration on how to use `ggvis`. As can be seen, ggvis can create graph in the same way ggplot does!

```{r}
p <- ggvis(mtcars, x = ~wt, y = ~ mpg)
layer_points(p)
```
Similar way to write basic ggvis

```{r}
layer_points(ggvis(mtcars, x = ~wt, y = ~ mpg))
```

Apply pipeline to write ggvis
```{r}
mtcars %>%
  ggvis(x = ~wt, y = ~mpg) %>%
  layer_points
```

We can add a theme on it

```{r}
mtcars %>% 
ggvis(x = ~wt, y = ~mpg) %>% 
layer_points() %>%
  add_axis("x", title = "Weight", ticks = 40,
    properties = axis_props(
      ticks = list(stroke = "red"),
      majorTicks = list(strokeWidth = 2),
      grid = list(stroke = "red"),
      labels = list(
        fill = "steelblue",
        angle = 50,
        fontSize = 14,
        align = "left",
        baseline = "middle",
        dx = 3
      ),
      title = list(fontSize = 16),
      axis = list(stroke = "#333", strokeWidth = 1.5)
    )
  )

```

## aesthetics
Try to adjust the parameters

```{r}
mtcars %>%
  ggvis(~mpg, ~disp, stroke = ~vs) %>%
  layer_points()
```

```{r}
mtcars %>% 
  ggvis(~mpg, ~disp, fill = ~vs) %>% 
  layer_points()
```

```{r}
mtcars %>% 
  ggvis(~mpg, ~disp, size = ~vs) %>% 
  layer_points()
```

```{r}
mtcars %>% 
  ggvis(~mpg, ~disp, shape = ~factor(cyl)) %>% 
  layer_points()
```

```{r}
mtcars %>% 
  ggvis(~wt, ~mpg, fill := "red",  stroke := "black") %>% 
  layer_points()
```

```{r}
mtcars %>% 
  ggvis(~wt, ~mpg, size := 300, opacity := 0.4) %>% 
  layer_points()

```

## Interaction

### basic interaction
```{r}
mtcars %>% ggvis(x = ~wt, y = ~mpg) %>%
  layer_points() %>%
  layer_smooths(
    span = input_slider(0.2, 1, value = 0.5, step = 0.05, label = "span")
  )
```
Next, we add some interaction strategies into it. We can add the following interaction into the graph:
- Change the size of the dot
- Change the opacity of the dot

```{r}
mtcars %>% 
   ggvis(~wt, ~mpg, 
         size := input_slider(10, 100),
         opacity := input_slider(0, 1)
   ) %>% 
   layer_points()
```

Add interaction strategies to a histograms
- Change the width of the histograms
- Change the center of the histograms

```{r}
mtcars %>% 
   ggvis(~wt) %>% 
   layer_histograms(
                    width =  input_slider(0, 2, step = 0.10, label = "width"),
                    center = input_slider(0, 2, step = 0.05, label = "center"))

```



Add tooltip

```{r}
mtcars %>%
   ggvis(~wt, ~mpg) %>% 
   layer_points()%>%
   add_tooltip(function(df) df$wt)

```

### Input option

#### input_select()

```{r}
mtcars %>% ggvis(x = ~wt) %>%
  layer_densities(
    adjust = input_slider(.1, 2, value = 1, step = .1, label = "Bandwidth adjustment"),
    kernel = input_select(
      c("Gaussian" = "gaussian", "Epanechnikov" = "epanechnikov",
        "Rectangular" = "rectangular", "Triangular" = "triangular",
        "Biweight" = "biweight", "Cosine" = "cosine", "Optcosine" = "optcosine"),
      label = "Kernel")
  )
```
#### Checkbox input
```{r}

mtcars %>% ggvis(x = ~wt, y = ~mpg) %>%
  layer_points(
    opacity := input_checkbox(label = "Semi-transparent",
                                    map = function(val) ifelse(val, .3, 1))) %>%
  layer_model_predictions(
    model = input_checkbox(label = "LOESS (curve) model fit",
                            map = function(val) ifelse(val, "loess", "lm")))
```

#### Text input
```{r}
mtcars %>% ggvis(x = ~wt, y = ~mpg) %>%
  layer_points(fill := input_text(label = "Point color", value = "red")) %>%
  layer_model_predictions(model = input_text(label = "Model type", value = "loess"))
```

#### Numeric input
```{r}
mtcars %>% ggvis(x = ~wt, y = ~mpg) %>%
  layer_points(size := input_numeric(value = 25, label = "Point size")) %>%
  layer_smooths(span = input_numeric(value = 0.5, label = "Interpolation points"))
```

#### Radio buttons
```{r}
mtcars %>% ggvis(x = ~wt, y = ~mpg) %>%
  layer_points() %>%
  layer_model_predictions(
    model = input_radiobuttons(c("LOESS" = "loess", "Linear" = "lm"),
      label = "Model type"),
    stroke := input_radiobuttons(c("Red" = "red", "Black" = "black"),
      label = "Line color")
  )
```


#### Checkbox group
```{r}
mtcars %>% ggvis(x = ~wt, y = ~mpg) %>%
  layer_points(
    fill := input_checkboxgroup(
      choices = c("Red" = "r", "Green" = "g", "Blue" = "b"),
      label = "Point color components",
      map = function(val) {
        rgb(0.8 * "r" %in% val, 0.8 * "g" %in% val, 0.8 * "b" %in% val)
      }
    )
  )
```

### Map

#### Example with value map function
```{r}
mtcars %>% ggvis(x = ~wt, y = ~mpg) %>%
  layer_points() %>%
  layer_model_predictions(model = "loess",
    model_args = list(n = input_select(
      choices = c("Two", "Six", "Eighty"),
      map = function(value) switch(value, Two = 2, Six = 6, Eighty = 80),
      label = "Number of points"
    ))
  )
```

```{r}

new_vals <- input_select(c("Set A" = "A", "Set B" = "B"),
  label = "Dynamically-generated column",
  map = function(value) {
    vals <- switch(value,
      "A" = rep(c("One", "Two")),
      "B" = c("First", "Second", "Third", "Fourth"))
    rep(vals, length = nrow(mtcars))
  })

mtcars %>% ggvis(x = ~wt, y = ~mpg, fill = new_vals) %>%
  layer_points()
```
```{r}
mtcars %>% ggvis(
    x = ~wt,
    y = ~mpg,
    fill = input_select(c("mpg", "wt"), map = as.name)
  ) %>%
  layer_points()
```


## Layers

### Simple layers

include primitives like points, lines and rectangles.

```{r}
mtcars %>% 
  ggvis(~wt, ~mpg) %>% 
  layer_points()
```

Paths and polygons

```{r}
df <- data.frame(x = 1:10, y = runif(10))
df %>% ggvis(~x, ~y) %>% layer_paths()
```

If you add `fill`, you will get a polygon
```{r}
t <- seq(0, 2 * pi, length = 100)
df <- data.frame(x = sin(t), y = cos(t))
df %>% ggvis(~x, ~y) %>% layer_paths(fill := "red")
```


```{r}
df <- data.frame(x = 1:10, y = runif(10))
df %>% ggvis(~x, ~y) %>% layer_ribbons()
```

Rectangle

```{r}
set.seed(1014)
df <- data.frame(x1 = runif(5), x2 = runif(5), y1 = runif(5), y2 = runif(5))
df %>% ggvis(~x1, ~y1, x2 = ~x2, y2 = ~y2, fillOpacity := 0.1) %>% layer_rects()
```

Text

```{r}
df <- data.frame(x = 3:1, y = c(1, 3, 2), label = c("a", "b", "c"))

df %>% ggvis(~x, ~y, text := ~label) %>% layer_text(fontSize := 50)
```

### Compound layers

```{r}
df <- data.frame(x = sin(t), y = cos(t))

df %>% ggvis(~x, ~y) %>% layer_paths()
```

```{r}
df %>% ggvis(~x, ~y) %>% layer_lines()
```

```{r}
mtcars %>% ggvis(~mpg) %>% layer_histograms()
```

```{r}
mtcars %>% ggvis(~wt, ~mpg) %>% layer_smooths()
```


