# Mschart

Huaqing Fang

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

The mschart package provides a framework for easily creating charts for ‘Microsoft PowerPoint’ documents. It has to be used with package officer that will produce the charts in new or existing PowerPoint or Word document

## Setup
install the mschart package, first install devtools. Then use ‘devtools’ to install maschart from Github, as shown in the code below.

```{r}
#install.packages("devtools")  ## install the package from github
#devtools::install_github("ardata-fr/mschart")
```



```{r}
library(mschart)
library(officer)
```

Use the Iris Dataset

```{r}
library(datasets)
data(iris)
```

## Create a line chart from a given dataframe

ms_linechart(data, x, y, group = NULL) creates a ms_linechart object where data is a given dataframe, x is the column name of x, y is the column name of y and group is the column name used to split data into series.

chart_ax_x(), chart_ax_y(...) defines settings for x and y axis

Note: rotation should be between -360 and 360 

Below are some commonly used parameters for num_fmt: \
"0": display the number with no decimal \
"0.00": display the number with two decimals \
"0\%\%": display as percentages\
"mm-dd-yy": display as date\

```{r}

linec <- mschart::ms_linechart(data = iris, x = "Sepal.Length",
                      y = "Sepal.Width", group = "Species")

linec <- mschart::chart_ax_y(linec, num_fmt = "0.00", rotation = -90)

```


## Produce the chart in a PowerPoint
The officer package lets R users manipulate Word (.docx) and PowerPoint (*.pptx) documents. Therefore, we can use this package to send the object as a chart.


```{r}
doc <- read_pptx() 

doc <- add_slide(doc, layout = "Title and Content", master = "Office Theme")

doc <- ph_with(doc, linec, location = ph_location(width = 4, height = 3, label = "hello"))

print(doc, target = "example.pptx") # print the new line chart into a target power point
                                    

```
To view it online: <a target="_blank" href="https://drive.google.com/file/d/1TpRVrJi7Q56OD9-oU4aBPY8q6oc6qc73/view?usp=sharing">office web viewer</a>

## View the chart by opening the PowerPoint

```{r}
browseURL("example.pptx") # use browseURL to open the target ppt
```

## Bar chart and other available charts
Other than linechart, we can also plot bar graphs, scatter plots and area graphs by calling:\
ms_barchart(...),\
ms_areachart(...),\
ms_scatterchart(...)


Note: read_pptx() opens a new powerpoint by default, to add graphs into an existing powerpoint, simply give the path of the existing document


```{r}
barc <- mschart::ms_barchart	(data = iris, x = "Sepal.Length",
                      y = "Sepal.Width", group = "Species")

barc <- mschart::chart_ax_y(barc, num_fmt = "0.00", rotation = -90)

doc <- read_pptx(path = "./example.pptx") 

doc <- add_slide(doc, layout = "Title and Content", master = "Office Theme")

doc <- ph_with(doc, barc, location = ph_location(width = 4, height = 3, label = "hello"))

print(doc, target = "example.pptx")
```
To view it online: <a target="_blank" href="https://drive.google.com/file/d/1PZlhMvCtTKLgTn0upSfhNm8_UTAmc0jP/view?usp=sharing">office web viewer</a>

## Write into a word document
simply use read_docx() and body_add_chart functions where the chart parameter can be any ms_chart object, that is one of [barchart, linechart, scatterchart, areachart]

```{R}
doc <- read_docx() 

doc <- body_add_chart(doc, chart = barc, style = "Normal")

print(doc, target = "example.docx")
```
To view it online: <a target="_blank" href="https://drive.google.com/file/d/1xxjUDtMafONITrFQ_NYZwqNZ4VjCNphj/view?usp=sharing">office web viewer</a>

## Customize Charts

Charts are generated with default values. Options are available to change charts properties.
Global chart settings can be specified with generic function chart_settings()
The following stacks the bars for each group.
```{r}
barc <- chart_settings(barc, grouping = "stacked", gap_width = 50, overlap = 100 )
```


Axis properties are available with functions chart_ax_x() and chart_ax_y()

```{r}
barc <- chart_ax_x(barc, cross_between = 'between', 
  major_tick_mark = "in", minor_tick_mark = "none")
barc <- chart_ax_y(barc, num_fmt = "0.00", rotation = -90)
```

Titles (main title, axis labels) can be set with function chart_labels():

```{r}
barc <- chart_labels(barc, title = "A main title", 
  xlab = "x axis title", ylab = "y axis title")
```


Series visual settings can be customized with functions:\
chart_data_fill(),\
chart_data_stroke(),\
chart_data_size(),\
chart_data_symbol(),\
chart_data_line_width() \
and chart_labels_text()
```{r}
barc <- chart_data_fill(barc,
  values = c(setosa = "#003C63", versicolor = "#ED1F24", virginica = "#F2AA00") )
barc <- chart_data_stroke(barc, values = "transparent" )
```

```{r}
doc <- read_docx() 

doc <- body_add_chart(doc, chart = barc, style = "Normal")

print(doc, target = "costumized.docx")
```
To view it online: <a target="_blank" href="https://drive.google.com/file/d/13W4jhh8ILay6kU2omQyyYyJLWixe4NtY/view?usp=sharing">office web viewer</a>
